---
title: "HumanBoneAging"
author: "Emine Guven"
date: "4 Ekim 2019"
output:
  html_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#GSE32719
#http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE32719


```{r }
rm(list=ls())

library(Biobase)
library(GEOquery)
library(limma)
library(foreach)
library(doMC)
library(gplots)
```

```{r comments out}
#setwd("C:/Users/User/AMH Dropbox/Emine Guven/gülsüm-burak-tez/analiz")
#CURRENT_DIR<-getwd()

#load platform data from bioconductor
#http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE32719
#GPL570	[HG-U133_Plus_2] Affymetrix Human Genome U133 Plus 2.0 Array
#https://bioconductor.org/packages/release/data/annotation/html/hgu133plus2.db.html
#Load series and platform data from GEO
```

```{r pull GEO data set}
#gset <- getGEO("GSE32719", GSEMatrix =TRUE);
#class(gset)
#str(gset);
#gset <- gset$GSE32719_series_matrix.txt.gz;
# make proper column names to match toptable 
# fvarLabels(gset) <- make.names(fvarLabels(gset))

# Save the object to a file to avoid pulling issues from the browser
#saveRDS(gset, file = "geneExpressionLevels.rds")
# Restore the object
gset<-readRDS(file = "geneExpressionLevels.rds")


```

##Check expression levels.

```{r pull exprs data}
ex= exprs(gset) ##gset içinden expres verisini çekip deðiþken ile tanýmladýk.
apply( ex, 2, median) 
means = apply( ex, 2, mean) ##Data içindeki kiþilerin tanýmlayýcý istatistiklerini kutu grafikte çizdirir. 
means
#pdf(paste("C:/Users/lenovo/Dropbox/analiz" , "boxplot.pdf", sep=''),width=5,height=5)
boxplot( ex )
#dev.off()
```

##Bu normalizasyon datanýn toplam yoðunluðuna göre gerçekleþir.
##Scale ile bir ölçekleme saðlanmaktadýr.
##for döngüsü için scale ile datadaki max. ortalama deðeri ile                                    ###tanýmladýk.

```{r normalization of expression levels}
scale = max(means)
for(i in 1:length(ex[1,])){
  ex[,i] = ex[,i] * scale/ means[i] }
apply( ex, 2, mean) / scale #all the means / scale(max.ort)
apply( ex, 2, median) /scale #dvide by scale to enable  normalization.
boxplot( ex )
```

```{r checking the data once more, results='hide'}
str(gset);
gset@phenoData@data; ##phenodata pull the data inside of it.(title, last update date, type, prtocol vb. bilgsini verir)
experimental_design = gset@phenoData@data;
```


```{r check the data}
experimental_design[1:3,] # check for the first individual to have look at quickly
# gset@phenoData@varMetadata
# experimental_design[, "source_name_ch1"][1:10]
experimental_design[1:3,  c("title", "source_name_ch1")]; ##add matrix column elements col.names.
unique( experimental_design$source_name_ch1 )
```


```{r annotation, results='hide'}
gpl <- annotation(gset) ##Biobase  annotation(gset) assigned it into  gpl variable 
platf <- getGEO(gpl, AnnotGPL=TRUE)
ncbifd <- data.frame(attr(dataTable(platf), "table"))
## platf  table[54675:21] matrix
###pull the annotated table from platf and assigned into ncbifd.
###the table includes info such as Gene ID,Gene title,Gene symbol 
```


```{r regression analysis to see age effect of the problem}
sample.names = unique(experimental_design$source_name_ch1) ###sperate groups into three O,M,Y
mylevels = c(1,2,3)
names(mylevels) = sample.names
gset.levels = mylevels[ experimental_design$source_name_ch1 ]
# p-values of ex
my.pvalues = numeric(length=length(ex[,1]))
# x2 = foreach(i=1:(3*2), .combine='cbind') %dopar% sqrt(i)

registerDoMC(cores=4) ##doMC package.
#my.pvalues = foreach( i = 1:100, .combine='rbind') %dopar% {
my.pvalues = foreach( i = 1:length(ex[,1]), .combine='rbind') %dopar% {
        m = lm( ex[i,] ~ gset.levels ) #gset.levels applies multiple regression analysis 
        ##lm function (linear model) regression analysis.
        sm = summary(m)
        pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)
}

row.names(my.pvalues) = row.names(gset@assayData$exprs)
```

```{r plot of pulled p-values}
summary(my.pvalues)
pdf(paste("grafikler/" ,"RegressionResultsPvalues.pdf",sep=''),width=8,height=5)
hist(my.pvalues,ylim=c(0,7000), col = "blue" )
dev.off()
```

```{r filter with Benjamini Hochberg statistics}
my.pvalues.BH = p.adjust(my.pvalues, "BH")#Benjamini, Hochberg
names(my.pvalues.BH) = row.names(gset@assayData$exprs)
pdf(paste("grafikler/", "RegressionResultsPvaluesBH.pdf",sep=''),width=8,height=5)
hist(my.pvalues.BH,ylim=c(0,8000), col = "green" )
dev.off()
summary(my.pvalues.BH)
```

```{r analyze BH output}
sig= my.pvalues.BH[my.pvalues.BH < 0.05]
sig= data.frame(sig)
sig$ID = row.names(sig)
sig2 = merge(sig, ncbifd, by="ID")
sig.genes = unique(sig2$Gene.symbol)


#BH criterion gene symbols
write.table(sig.genes, "__sig.genes.tsv", sep="\t", quote=F, row.names=F, col.names=F  )
write.csv(sig.genes,"sig.genes.csv",sep="\t" )

#backgorund genes for the GO analysis
background.genes = unique( ncbifd$Gene.symbol)
write.table(background.genes, "__background.genes.tsv", sep="\t", quote=F, row.names=F, col.names=F  )
write.csv(background.genes,"background.genes.csv",sep="\t" )

# for the GO analysis

sigGO<-data.frame("geneSymbol"=sig2$Gene.symbol,"geneFunc"=sig2$GO.Function,
                  "geneFuncId"=sig2$GO.Function.ID,"geneComponent"=sig2$GO.Component,                                        "geneComponentId"=sig2$GO.Component.ID, 
                  "geneProcess"=sig2$GO.Process, "geneProcessID"=sig2$GO.Process.ID)

write.table(sigGO, "_sigGO.csv",sep="\t")
write.csv(sigGO,"sigGO.csv",sep="\t" )

```

```{r more figures of aging related analysis }
plot(experimental_design$source_name_ch1, ylab="Number of Samples",break=)
plot(experimental_design$characteristics_ch1.2)

sml <- c("Y","Y","Y","Y","Y","Y","Y","Y","Y","Y","Y","Y","Y","Y",
         "M","M","M","M","M","O","O","O","O","O","O","O","O")

fl <- as.factor(sml)
labels <- c("MIDDLE","OLD","YOUNG")

palette(c("yellow","green","#blue"))
title <- paste ("GSE32719", '/', gpl, " SELECTED SAMPLES", sep ='')
#pdf(paste("C:/Users/User/AMH Dropbox/Emine Guven/gülsüm-burak-tez/analiz/grafikler/", "boxplot.pdf",sep=''),width=8,height=5)
boxplot(ex, boxwex=0.75, notch=T, main=title, outline=FALSE, las=2, col=fl)
legend("topright", labels, fill=palette(), bty="n")
#dev.off()

hist(ex, ylim=c(0,250000),col = "blue", main="GSE32719 - Histogram")
log = log2(ex)
hist(log, col = "gray", main="GSE32719 (log2) - Histogram")

hc = hclust(as.dist(1-cor(ex)))
plot(hc, main="GSE32719 - Hierarchical Clustering")


