---
title: "20160117 GSE32719 FL_HHMI15 Aging in HSC" 
author: "h qin"
date: "January 17, 2016"
output: html_document
---

GSE32719
http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE32719

```{r, message=FALSE}
rm(list=ls())
setwd("~/github/FL_HHMI15/project1.GSE32719")
library(Biobase)
library(GEOquery)
library(limma)
library(foreach)
library(doMC)
```

load platform data from bioconductor
http://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE32719
GPL570	[HG-U133_Plus_2] Affymetrix Human Genome U133 Plus 2.0 Array
https://bioconductor.org/packages/release/data/annotation/html/hgu133plus2.db.html


Load series and platform data from GEO
# gset <- getGEO("GSE32719", GSEMatrix =FALSE)
gset


 gset <- gset$GSE32719_series_matrix.txt.gz
# make proper column names to match toptable 
# fvarLabels(gset) <- make.names(fvarLabels(gset))




ex= exprs(gset)
apply( ex, 2, median)
means = apply( ex, 2, mean)
means
boxplot( ex )


scale = max(means)
for(i in 1:length(ex[1,])){
  ex[,i] = ex[,i] * scale/ means[i]
}
apply( ex, 2, mean) / scale
apply( ex, 2, median) /scale
boxplot( ex )



str(gset)
#gset@phenoData@data
experimental_design = gset@phenoData@data
experimental_design[1:3,]
# gset@phenoData@varMetadata
# experimental_design[, "source_name_ch1"][1:10]
experimental_design[1:3,  c("title", "source_name_ch1")]
unique( experimental_design$source_name_ch1 



gpl <- annotation(gset)
platf <- getGEO(gpl, AnnotGPL=TRUE)
ncbifd <- data.frame(attr(dataTable(platf), "table"))




sample.names = unique(experimental_design$source_name_ch1)
mylevels = c(1,2,3)
names(mylevels) = sample.names
gset.levels = mylevels[ experimental_design$source_name_ch1 ]
my.pvalues = numeric(length=length(ex[,1]))
# x2 = foreach(i=1:(3*2), .combine='cbind') %dopar% sqrt(i)

registerDoMC(cores=4)
#my.pvalues = foreach( i = 1:100, .combine='rbind') %dopar% {
my.pvalues = foreach( i = 1:length(ex[,1]), .combine='rbind') %dopar% {
  m = lm( ex[i,] ~ gset.levels )
  sm = summary(m)
  pf(sm$fstatistic[1], sm$fstatistic[2], sm$fstatistic[3], lower.tail = FALSE)
}
row.names(my.pvalues) = row.names(gset@assayData$exprs)



hist(my.pvalues)
summary(my.pvalues)
my.pvalues.BH = p.adjust(my.pvalues, "BH")
names(my.pvalues.BH) = row.names(gset@assayData$exprs)
hist(my.pvalues.BH)
summary(my.pvalues.BH)
sig= my.pvalues.BH[my.pvalues.BH < 0.05]

sig= data.frame(sig)
sig$ID = row.names(sig)
#ncbifd[ match(names(sig), as.character( ncbifd$ID) ),  ]
sig2 = merge(sig, ncbifd, by="ID")
```

GO analysis
```{r}
sig.genes = unique(sig2$Gene.symbol)
#sig.geneIDs = unique(sig2$Gene.ID)
write.table(sig.genes, "__sig.genes.tsv", sep="\t", quote=F, row.names=F, col.names=F  )

background.genes = unique( ncbifd$Gene.symbol)
write.table(background.genes, "__background.genes.tsv", sep="\t", quote=F, row.names=F, col.names=F  )

#library(topGO)
```
GOrilla show 

Todo: Load bioconductor annotation
```{r, message=FALSE }
# source("https://bioconductor.org/biocLite.R")
# biocLite("hgu133plus2.db")
# library("hgu133plus2.db")
# help(package="hgu133plus2.db")
# ls("package:hgu133plus2.db")
# x <- hgu133plus2GENENAME
# x <- hgu133plus2SYMBOL
# mapped_probes <- mappedkeys(x)
# xx <- as.list(x[mapped_probes])
# xx[1:5]
```
